LHC abort gap monitor electronics upgrade

### 2024-04-21 paper.tex
chapters filled with drafts

\section{Introduction}
The maximum beam energy in~LHC at flat-top distributed in 2808 bunches can
reach \SI{350}{MJ}, while having its average transverse size smaller than
\SI{1}{mm}[LHCreport]. Such energy density can cause damage to the machine. 
Inervention with conductive surface is enough to melt metallic surfaces and
quenching a superconductiv magnet. The Beam dump system safely removes beam
from the vacuum chamber. Its kicker magnets need \SI{3}{\micro\second} to ramp
up their magnetic field and diverge particles towards beam dumps. The particles
inside AG due to inperfections get insufficient kick and deflect towards the
vacuum chamber. At certain energy accumulated in AG, the magnet can quench.
Therefore continuous monitoring of the AG population is required. This is the
purpose of the AGM~[abortgapmonitor].\\ 
This paper reviews upgrade of the current AGM system installed in the LHC since
2010. It describes first general system architecture and analog electronics
chain of the current implementation. Discusses the temperature dependence,
timing instabilities and compares the current and upgraded system.

\section{System architecture}
Abort gap monitoring methodology is non-invasive, because it exploits
synchrotron radiation light (SL). This phenomena appears, when a particle is
losing it's energy on a circular trajectory. The energy is emitted in the form
of light tangentially to the curved beam. The spectrum of the SL is
proportional to the energy of the particles in the beam [REF]. At the flat-top
energy (6.8TeV) the light is in visible range and can be measured by a
photomultiplier (PMT), which converts light to electric current.

\subsection{Working principle}
The~AGM uses the SR
spontaneously emitted by the beam to quantify the fraction of particles present
in the AG.

System operational principle is depicted in the Fig.~\ref{fig:bsra_chain}.

The signal is collected in the point 4, where it is measured by the PMT. The
signal is converted and amplified depending on the voltage at the anode.

The Emitted SL is collected by the PMT. It converts light to the electrict
signal. This goes to through amplifier to the integrator, where it is decimated
by factor 10. The system measure window is \SI{6.4}{\micro\second} long, which
consist of 64 values after decimation process. Each of these values goes to the
Exponential moving averager. This type of filter has a feedback remmembering
all samples from system start.  
\begin{figure}[!tbh]
    \centering
    \includegraphics[width=\columnwidth]{bsra_chain.pdf}
    \caption{BSRA Chain}
    \label{fig:bsra_chain}
\end{figure}
nteaohutnha tnaheusnthae nsatheusnha thoaeut hh antuhentae u tahuenth tahuenth
nteaohutnha tnaheusnthae nsatheusnha thoaeut hh antuhentae u tahuenth tahuenth
nteaohutnha tnaheusnthae nsatheusnha thoaeut hh antuhentae u tahuenth tahuenth
nteaohutnha tnaheusnthae nsatheusnha thoaeut hh antuhentae u tahuenth tahuenth
nteaohutnha tnaheusnthae nsatheusnha thoaeut hh antuhentae u tahuenth tahuenth
nteaohutnha tnaheusnthae nsatheusnha thoaeut hh antuhentae u tahuenth tahuenth
nteaohutnha tnaheusnthae nsatheusnha thoaeut hh antuhentae u tahuenth tahuenth
nteaohutnha tnaheusnthae nsatheusnha thoaeut hh antuhentae u tahuenth tahuenth
nteaohutnha tnaheusnthae nsatheusnha thoaeut hh antuhentae u tahuenth tahuenth
nteaohutnha tnaheusnthae nsatheusnha thoaeut hh antuhentae u tahuenth tahuenth
nteaohutnha tnaheusnthae nsatheusnha thoaeut hh antuhentae u tahuenth tahuenth
nteaohutnha tnaheusnthae nsatheusnha thoaeut hh antuhentae u tahuenth tahuenth
nteaohutnha tnaheusnthae nsatheusnha thoaeut hh antuhentae u tahuenth tahuenth
nteaohutnha tnaheusnthae nsatheusnha thoaeut hh antuhentae u tahuenth tahuenth

\subsection{Current Solution}
nteaohutnha tnaheusnthae nsatheusnha thoaeut hh antuhentae u tahuenth tahuenth
nteaohutnha tnaheusnthae nsatheusnha thoaeut hh antuhentae u tahuenth tahuenth
nteaohutnha tnaheusnthae nsatheusnha thoaeut hh antuhentae u tahuenth tahuenth
\begin{figure}[!tbh]
    \centering
    \includegraphics[width=\columnwidth]{existing_system_processing.pdf}
    \caption{Current system block diagram}
    \label{fig:current_system_processing}
\end{figure}
nteaohutnha tnaheusnthae nsatheusnha thoaeut hh antuhentae u tahuenth tahuenth
nteaohutnha tnaheusnthae nsatheusnha thoaeut hh antuhentae u tahuenth tahuenth
nteaohutnha tnaheusnthae nsatheusnha thoaeut hh antuhentae u tahuenth tahuenth
nteaohutnha tnaheusnthae nsatheusnha thoaeut hh antuhentae u tahuenth tahuenth
nteaohutnha tnaheusnthae nsatheusnha thoaeut hh antuhentae u tahuenth tahuenth
nteaohutnha tnaheusnthae nsatheusnha thoaeut hh antuhentae u tahuenth tahuenth
nteaohutnha tnaheusnthae nsatheusnha thoaeut hh antuhentae u tahuenth tahuenth
nteaohutnha tnaheusnthae nsatheusnha thoaeut hh antuhentae u tahuenth tahuenth
nteaohutnha tnaheusnthae nsatheusnha thoaeut hh antuhentae u tahuenth tahuenth
nteaohutnha tnaheusnthae nsatheusnha thoaeut hh antuhentae u tahuenth tahuenth

\subsection{Constraints and Limitations}
% This relies on an analogue
% integrator ASIC and a 40 MHz analogue to digital converter (ADC) for particle
% population measurements. Recognizing the obsolescence of this analogue
% solution, the paper introduces a transition to a digital signal processing
% approach. Working directly in the digital domain not only offers more
% scalability but also better determinism and reliability.
nteaohutnha tnaheusnthae nsatheusnha thoaeut hh antuhentae u tahuenth tahuenth
nteaohutnha tnaheusnthae nsatheusnha thoaeut hh antuhentae u tahuenth tahuenth
nteaohutnha tnaheusnthae nsatheusnha thoaeut hh antuhentae u tahuenth tahuenth
nteaohutnha tnaheusnthae nsatheusnha thoaeut hh antuhentae u tahuenth tahuenth
nteaohutnha tnaheusnthae nsatheusnha thoaeut hh antuhentae u tahuenth tahuenth
nteaohutnha tnaheusnthae nsatheusnha thoaeut hh antuhentae u tahuenth tahuenth
nteaohutnha tnaheusnthae nsatheusnha thoaeut hh antuhentae u tahuenth tahuenth

\section{System Upgrade}
nteaohutnha tnaheusnthae nsatheusnha thoaeut hh antuhentae u tahuenth tahuenth
nteaohutnha tnaheusnthae nsatheusnha thoaeut hh antuhentae u tahuenth tahuenth
nteaohutnha tnaheusnthae nsatheusnha thoaeut hh antuhentae u tahuenth tahuenth
nteaohutnha tnaheusnthae nsatheusnha thoaeut hh antuhentae u tahuenth tahuenth
nteaohutnha tnaheusnthae nsatheusnha thoaeut hh antuhentae u tahuenth tahuenth
nteaohutnha tnaheusnthae nsatheusnha thoaeut hh antuhentae u tahuenth tahuenth
nteaohutnha tnaheusnthae nsatheusnha thoaeut hh antuhentae u tahuenth tahuenth
%Utilization of Analogue Integrator ASIC
%40 MHz Analogue-to-Digital Converter (ADC)
%Operation Since 2010
%Reasons for Obsolescence
%Digital Signal Processing Approach why and outcomes

\subsection{Acquisition System}
nteaohutnha tnaheusnthae nsatheusnha thoaeut hh antuhentae u tahuenth tahuenth
nteaohutnha tnaheusnthae nsatheusnha thoaeut hh antuhentae u tahuenth tahuenth
nteaohutnha tnaheusnthae nsatheusnha thoaeut hh antuhentae u tahuenth tahuenth
\begin{figure}[!tbh]
    \centering
    \includegraphics[width=\columnwidth]{bsra_daq.pdf}
    \caption{Ackquisition system schematics}
    \label{fig:bsra_daq}
\end{figure}
nteaohutnha tnaheusnthae nsatheusnha thoaeut hh antuhentae u tahuenth tahuenth
nteaohutnha tnaheusnthae nsatheusnha thoaeut hh antuhentae u tahuenth tahuenth
nteaohutnha tnaheusnthae nsatheusnha thoaeut hh antuhentae u tahuenth tahuenth
nteaohutnha tnaheusnthae nsatheusnha thoaeut hh antuhentae u tahuenth tahuenth
nteaohutnha tnaheusnthae nsatheusnha thoaeut hh antuhentae u tahuenth tahuenth
nteaohutnha tnaheusnthae nsatheusnha thoaeut hh antuhentae u tahuenth tahuenth
nteaohutnha tnaheusnthae nsatheusnha thoaeut hh antuhentae u tahuenth tahuenth
nteaohutnha tnaheusnthae nsatheusnha thoaeut hh antuhentae u tahuenth tahuenth
nteaohutnha tnaheusnthae nsatheusnha thoaeut hh antuhentae u tahuenth tahuenth
nteaohutnha tnaheusnthae nsatheusnha thoaeut hh antuhentae u tahuenth tahuenth
nteaohutnha tnaheusnthae nsatheusnha thoaeut hh antuhentae u tahuenth tahuenth
nteaohutnha tnaheusnthae nsatheusnha thoaeut hh antuhentae u tahuenth tahuenth
nteaohutnha tnaheusnthae nsatheusnha thoaeut hh antuhentae u tahuenth tahuenth
nteaohutnha tnaheusnthae nsatheusnha thoaeut hh antuhentae u tahuenth tahuenth
nteaohutnha tnaheusnthae nsatheusnha thoaeut hh antuhentae u tahuenth tahuenth
nteaohutnha tnaheusnthae nsatheusnha thoaeut hh antuhentae u tahuenth tahuenth

\subsection{Low band Amplifier}
% \input{tables/system_comparison.tex} 
%compare the changes and comment the expected changes
nteaohutnha tnaheusnthae nsatheusnha thoaeut hh antuhentae u tahuenth tahuenth
nteaohutnha tnaheusnthae nsatheusnha thoaeut hh antuhentae u tahuenth tahuenth
nteaohutnha tnaheusnthae nsatheusnha thoaeut hh antuhentae u tahuenth tahuenth
nteaohutnha tnaheusnthae nsatheusnha thoaeut hh antuhentae u tahuenth tahuenth
nteaohutnha tnaheusnthae nsatheusnha thoaeut hh antuhentae u tahuenth tahuenth
nteaohutnha tnaheusnthae nsatheusnha thoaeut hh antuhentae u tahuenth tahuenth
nteaohutnha tnaheusnthae nsatheusnha thoaeut hh antuhentae u tahuenth tahuenth
nteaohutnha tnaheusnthae nsatheusnha thoaeut hh antuhentae u tahuenth tahuenth
nteaohutnha tnaheusnthae nsatheusnha thoaeut hh antuhentae u tahuenth tahuenth
\begin{figure}[!tbh]
    \centering
    \includegraphics[width=\columnwidth]{amp_schematics.pdf}
    \caption{upgraded amplifier}
    \label{fig:amp_schematics}
\end{figure}
nteaohutnha tnaheusnthae nsatheusnha thoaeut hh antuhentae u tahuenth tahuenth
nteaohutnha tnaheusnthae nsatheusnha thoaeut hh antuhentae u tahuenth tahuenth
nteaohutnha tnaheusnthae nsatheusnha thoaeut hh antuhentae u tahuenth tahuenth
nteaohutnha tnaheusnthae nsatheusnha thoaeut hh antuhentae u tahuenth tahuenth
nteaohutnha tnaheusnthae nsatheusnha thoaeut hh antuhentae u tahuenth tahuenth
nteaohutnha tnaheusnthae nsatheusnha thoaeut hh antuhentae u tahuenth tahuenth
nteaohutnha tnaheusnthae nsatheusnha thoaeut hh antuhentae u tahuenth tahuenth
nteaohutnha tnaheusnthae nsatheusnha thoaeut hh antuhentae u tahuenth tahuenth
nteaohutnha tnaheusnthae nsatheusnha thoaeut hh antuhentae u tahuenth tahuenth
nteaohutnha tnaheusnthae nsatheusnha thoaeut hh antuhentae u tahuenth tahuenth
nteaohutnha tnaheusnthae nsatheusnha thoaeut hh antuhentae u tahuenth tahuenth
nteaohutnha tnaheusnthae nsatheusnha thoaeut hh antuhentae u tahuenth tahuenth
nteaohutnha tnaheusnthae nsatheusnha thoaeut hh antuhentae u tahuenth tahuenth
nteaohutnha tnaheusnthae nsatheusnha thoaeut hh antuhentae u tahuenth tahuenth

\subsection{Installation in LHC}
nteaohutnha tnaheusnthae nsatheusnha thoaeut hh antuhentae u tahuenth tahuenth
nteaohutnha tnaheusnthae nsatheusnha thoaeut hh antuhentae u tahuenth tahuenth
nteaohutnha tnaheusnthae nsatheusnha thoaeut hh antuhentae u tahuenth tahuenth
nteaohutnha tnaheusnthae nsatheusnha thoaeut hh antuhentae u tahuenth tahuenth
nteaohutnha tnaheusnthae nsatheusnha thoaeut hh antuhentae u tahuenth tahuenth
nteaohutnha tnaheusnthae nsatheusnha thoaeut hh antuhentae u tahuenth tahuenth
nteaohutnha tnaheusnthae nsatheusnha thoaeut hh antuhentae u tahuenth tahuenth
nteaohutnha tnaheusnthae nsatheusnha thoaeut hh antuhentae u tahuenth tahuenth
nteaohutnha tnaheusnthae nsatheusnha thoaeut hh antuhentae u tahuenth tahuenth
\begin{figure}[!tbh]
    \centering
    \includegraphics[width=\columnwidth]{installation.jpg}
    \caption{installation picture}
    \label{fig:installation}
\end{figure}
nteaohutnha tnaheusnthae nsatheusnha thoaeut hh antuhentae u tahuenth tahuenth
nteaohutnha tnaheusnthae nsatheusnha thoaeut hh antuhentae u tahuenth tahuenth
nteaohutnha tnaheusnthae nsatheusnha thoaeut hh antuhentae u tahuenth tahuenth
nteaohutnha tnaheusnthae nsatheusnha thoaeut hh antuhentae u tahuenth tahuenth
nteaohutnha tnaheusnthae nsatheusnha thoaeut hh antuhentae u tahuenth tahuenth

\section{Control Interface}
nteaohutnha tnaheusnthae nsatheusnha thoaeut hh antuhentae u tahuenth tahuenth
nteaohutnha tnaheusnthae nsatheusnha thoaeut hh antuhentae u tahuenth tahuenth
\begin{figure}[!tbh]
    \centering
    \includegraphics[width=\columnwidth]{control_interface.jpg}
    \caption{installation picture}
    \label{fig:control_interface}
\end{figure}
nteaohutnha tnaheusnthae nsatheusnha thoaeut hh antuhentae u tahuenth tahuenth
nteaohutnha tnaheusnthae nsatheusnha thoaeut hh antuhentae u tahuenth tahuenth
nteaohutnha tnaheusnthae nsatheusnha thoaeut hh antuhentae u tahuenth tahuenth
nteaohutnha tnaheusnthae nsatheusnha thoaeut hh antuhentae u tahuenth tahuenth
nteaohutnha tnaheusnthae nsatheusnha thoaeut hh antuhentae u tahuenth tahuenth
nteaohutnha tnaheusnthae nsatheusnha thoaeut hh antuhentae u tahuenth tahuenth
nteaohutnha tnaheusnthae nsatheusnha thoaeut hh antuhentae u tahuenth tahuenth
nteaohutnha tnaheusnthae nsatheusnha thoaeut hh antuhentae u tahuenth tahuenth
nteaohutnha tnaheusnthae nsatheusnha thoaeut hh antuhentae u tahuenth tahuenth
nteaohutnha tnaheusnthae nsatheusnha thoaeut hh antuhentae u tahuenth tahuenth

\section{Measurements}
nteaohutnha tnaheusnthae nsatheusnha thoaeut hh antuhentae u tahuenth tahuenth
nteaohutnha tnaheusnthae nsatheusnha thoaeut hh antuhentae u tahuenth tahuenth
nteaohutnha tnaheusnthae nsatheusnha thoaeut hh antuhentae u tahuenth tahuenth
nteaohutnha tnaheusnthae nsatheusnha thoaeut hh antuhentae u tahuenth tahuenth
\begin{figure}[!htb]
    \begin{center}
        \scalebox{0.54}{\input{plots/ions.pgf}}
        \caption{Comparison of AGMs}
        \label{fig:comparison_chart}
    \end{center}
\end{figure}
nteaohutnha tnaheusnthae nsatheusnha thoaeut hh antuhentae u tahuenth tahuenth
nteaohutnha tnaheusnthae nsatheusnha thoaeut hh antuhentae u tahuenth tahuenth
nteaohutnha tnaheusnthae nsatheusnha thoaeut hh antuhentae u tahuenth tahuenth
nteaohutnha tnaheusnthae nsatheusnha thoaeut hh antuhentae u tahuenth tahuenth
nteaohutnha tnaheusnthae nsatheusnha thoaeut hh antuhentae u tahuenth tahuenth

\section{Discussion}
nteaohutnha tnaheusnthae nsatheusnha thoaeut hh antuhentae u tahuenth tahuenth
nteaohutnha tnaheusnthae nsatheusnha thoaeut hh antuhentae u tahuenth tahuenth
nteaohutnha tnaheusnthae nsatheusnha thoaeut hh antuhentae u tahuenth tahuenth
nteaohutnha tnaheusnthae nsatheusnha thoaeut hh antuhentae u tahuenth tahuenth
nteaohutnha tnaheusnthae nsatheusnha thoaeut hh antuhentae u tahuenth tahuenth
nteaohutnha tnaheusnthae nsatheusnha thoaeut hh antuhentae u tahuenth tahuenth
nteaohutnha tnaheusnthae nsatheusnha thoaeut hh antuhentae u tahuenth tahuenth

\section{Conclusion}
nteaohutnha tnaheusnthae nsatheusnha thoaeut hh antuhentae u tahuenth tahuenth
nteaohutnha tnaheusnthae nsatheusnha thoaeut hh antuhentae u tahuenth tahuenth
nteaohutnha tnaheusnthae nsatheusnha thoaeut hh antuhentae u tahuenth tahuenth
nteaohutnha tnaheusnthae nsatheusnha thoaeut hh antuhentae u tahuenth tahuenth
nteaohutnha tnaheusnthae nsatheusnha thoaeut hh antuhentae u tahuenth tahuenth
nteaohutnha tnaheusnthae nsatheusnha thoaeut hh antuhentae u tahuenth tahuenth
nteaohutnha tnaheusnthae nsatheusnha thoaeut hh antuhentae u tahuenth tahuenth
nteaohutnha tnaheusnthae nsatheusnha thoaeut hh antuhentae u tahuenth tahuenth
nteaohutnha tnaheusnthae nsatheusnha thoaeut hh antuhentae u tahuenth tahuenth
nteaohutnha tnaheusnthae nsatheusnha thoaeut hh antuhentae u tahuenth tahuenth
nteaohutnha tnaheusnthae nsatheusnha thoaeut hh antuhentae u tahuenth tahuenth
nteaohutnha tnaheusnthae nsatheusnha thoaeut hh antuhentae u tahuenth tahuenth

% \input{acknowledgements}
% \input{appendix}
\input{bibliography}

%%%\include{annexes-A4}

>>>>

\section{Introduction}
The LHC can store a maximum beam energy of up to 350 MJ across 2808 bunches
during its flat-top phase, with an average transverse beam size under 1 mm
(LHCreport). This high energy density poses a risk of machine damage, as
conductive surface intervention can melt metallic surfaces and quench
superconducting magnets. The Beam Dump system, utilizing kicker magnets with a
3-microsecond ramp-up time to diverting particles, takes care of removing the beam
from the vacuum chamber. It exploits a gap in the beam called abort gap (AG),
to ramp up the magnets and divert the beam towards beam dumps.
Imperfections in the accelerator may result in accumulation of particles inside
AG. This results in insufficient kicks, causing particles to deflect
towards the vacuum chamber and potentially leading to magnet quenching. This
why the AG needs to be continuously monitored by the abort gap monitor (AGM).

This paper reviews the upgrade of the current AGM system in the LHC since 2010,
focusing on system architecture, analog electronics chain, temperature
dependence, timing instabilities, and a comparison with the upgraded system.

The maximum beam energy in~LHC at flat-top distributed in 2808 bunches can
reach \SI{350}{MJ}, while having its average transverse size smaller than
\SI{1}{mm}[LHCreport]. Such energy density can cause damage to the machine. 
Inervention with conductive surface is enough to melt metallic surfaces and
quenching a superconductiv magnet. The Beam dump system safely removes beam
from the vacuum chamber. Its kicker magnets need \SI{3}{\micro\second} to ramp
up their magnetic field and diverge particles towards beam dumps. The particles
inside AG due to inperfections get insufficient kick and deflect towards the
vacuum chamber. At certain energy accumulated in AG, the magnet can quench.
Therefore continuous monitoring of the AG population is required. This is the
purpose of the AGM~[abortgapmonitor].\\ 
This paper reviews upgrade of the current AGM system installed in the LHC since
2010. It describes first general system architecture and analog electronics
chain of the current implementation. Discusses the temperature dependence,
timing instabilities and compares the current and upgraded system.

\section{System architecture}
System is assambled from three major parts. First part is the optical font-end
including preamplifier placed under the vacuum chamber. Second part is a
gateware implemented on FPGA platform placed outside of the radiation area. The
last part is the software reading out measured data and providing interface to
interact with the gateware. The system architecture is depicted in
Fig.~\ref{fig:bsra_chain}, where each part is enclosed in a frame with a
number.
\begin{figure}[!tbh]
    \centering
    \includegraphics[width=\columnwidth]{bsra_chain.pdf}
    \caption{BSRA Chain}
    \label{fig:bsra_chain}
\end{figure}

\subsection{Working principle}
Abort gap monitoring is non-invasive beam measurement with high dynamic range.
It exploits synchrotron radiation light (SL) present, when trajectory of a
particle is bend due to magnetic forces. The particle loses it's energy in a
form of light tangentally radiated to its trajectory. Any change in the light
spectrum reflects the change in the beam [REF]. 
The SL spectrum at the flat-top energy (6.8TeV) is in visible range, which
is measured by a optical front-end. This consists of photomultiplier (PMT),
converting SL to electric current, filters, and pre-amplifier to increase
magnitude before sending the signal \SI{100}{m} far away data acquisition
system (DAQ). The PMT is gate, that it opens just at the moment of abort gap.
\\
When SL is converted it is amplified and send to the DAQ. The system is
synchronous with the RF timing, that it gets information about bunch position
and turn start. The signal goes to through amplifier to the integrator, where
it is decimated by factor 10. The system is gated. Its measure window is
\SI{6.4}{\micro\second} long, but the gate opens just for the time of the abort
gap. The decimate the data to \SI{100}{ns} sample period. In other words system
measures 64 values, but just 30 have valid value. Each of these values goes to
the Exponential moving averager. This type of filter has a feedback, so the
its output reflects the evolution of the charge in the abort gap.

\subsection{Current Solution}
The System is placed at the point 4 sharing the optical line with the BSRT
[REF]. It converts incoming light through MCP-PMT HAMAMATSU R5916U-50 into
electric current. The current is converted to voltage by transimpedance
amplifier HAMAMATSU C5594 with static gain \SI{36}{dBV}, bandwith \SI{1.5}{GHz}
and the low cutoff frequency 50kHz.
The amplified signal is send to the switching analog integrators inside ASIC
lhcb2002b mounted to the IBMS FMC card. The card is connected to the DAB64x
deploying FPGA Altera Stratix EP1S400. 
\begin{figure}[!tbh]
    \centering
    \includegraphics[width=\columnwidth]{existing_system_processing.pdf}
    \caption{Current system block diagram}
    \label{fig:current_system_processing}
\end{figure}
The current system schematic is in Fig.~\ref{fig:current_system_processing}. It
uses preintegration in the analog integrators behind the pre-amplifier. One
outputs a value while the other one is measuring every \SI{25}{ns}. Inside the
gateware the samples are summed to get 100ns integrals. Each analog integrator
has different gain, offset and is not temperature invariant. Therefore
different coefficients are applied at this stage as well. 
The system provides the two sets of data. It measures the evolution of average
and standard deviation. The 2 x 64 values are stored in the Block RAM and
fetched by the PC with running FESA server.

\subsection{Constraints and Limitations}
The system bottle neck is the analog integration. There is one installation per
beam each having IBMS with different gain, offset, correction coefficients and
40 MHz ADCs. When the IBMS is replaced, it needs to be recalibrated. The ASIC
has not deterministic time stamps, so dead-times lengths do not match. This is
causing artefacts at the opening and closing of the gate in the output plot.
The dynamic range can be controled by the optical filters not via amplifier,
because of static gain. Finally the DAB64x is obsolate platform, which
complicates its maintenance.

\section{System Upgrade}
The motivation to create a new Abort Gap Monitor is initiated by the problems
associated with the analog integrators in lhcb2002b. These cause measurements
errors and make the AGM difficult to calibrate[REF]. And the second driver to
redesign the system was the platform obsolance reducing simplicity of
maintenance. \\ 
As a platform the new VFC-HD Carrier is used with additional RTM board
providing driver for the gate and control of the built-in test LED. The last
major change was of the shelf 500MSPS 14-bit ADC reducing the maintanence
efford.

\subsection{Low band Amplifier}
The amplifier installed close to the MCP-PMT was custom designed. The purpose
is different light intensity of ion and proton beams to obtain higher dynamic
range. \\ 
% A = U^2/(i^2 * R^2) => A_dB = 20*log10(U/(i*R))
In the Fig. \ref{fig:amp_schematics} is shown amplifier's block scheme. All the
settings is controlled via UART interface. It provides changing the variable
current-to-voltage conversion factor from \SI{5.8}{mV/\micro A} -
\SI{1.1}{V/\micro A}. This corresponds to the Voltage gain from \SI{41}{dB} to
\SI{87}{dB}. The ADC has a no zero offset, so the signal can be shifted in the
amplifier before post-processing.\\
The amplifier features also alarm signal generator. This serves to turn of the
gate, when the light impinging is to intense and it could damage the
multi-channel plate inside the PMT. The alarm signal is send to the gateware,
where it is handled. The alarm feature has a variable threshold set.   
\begin{figure}[!tbh]
    \centering
    \includegraphics[width=\columnwidth]{amp_schematics.pdf}
    \caption{upgraded amplifier}
    \label{fig:amp_schematics}
\end{figure}
The amplifier design has three switches. First switch can set pre-integrator,
to filter out spikes caused by dark counts of the PMT. Second switch can
activate either manual or gate driven integration of the acquisition window.  
Last switch activates low pass filtering with cutoff frequency of 40MHz.

\subsection{Acquisition system}
The DAQ was completly redesigned, because of the new platform and additional HW
components. The Fig. \ref{fig:bsra_daq} shows, how the signal is processed an
published from the DAQ.
\begin{figure}[!tbh]
    \centering
    \includegraphics[width=\columnwidth]{bsra_daq.pdf}
    \caption{Ackquisition system schematics}
    \label{fig:bsra_daq}
\end{figure}
The ADC takes amplified data from MCP-PMT and samples them with 500MSPS
asynchronously to the LHC timing. To calculate the integrals, the system tags
each of the sample. This is done, that tagging block resamples LHC timing by
500MSPS and creates bunch and turn clock vector masks. Those store the
information about both clocks rising edge phase. The 14-bit data from ADC are
left aligned to 16-bit values. The tagger sets their two LSBs to one, when the
ADC sample falls into the edge range. \\
The ADC provides four frames of four 16-bit samples every \SI{8}{\nano s} each
belonging to one ADC channel. The tag can appear in any of the four values. The
system stores data from region of interest, using FIFO. The synchronizer counts
4 turns and then generates a trigger for FIFO. Stored data are then processed as
a stream of values passed to the integrator. This starts summing up values,
when it hits the tag in the data marking the start of the bunch. It creates 64
values offseted by the user when needed. This correct the baseline. The values
are then sent to the exponential moving averager. The result is then stored in
the block ram available on to read from VME64x bus.

\subsection{Installation in LHC}
The instrument front and Acquistion system is located in Point 4 near RF
system. The new installation is shown in Fig. \ref{fig:installation}
highlighted with violet frame. On its right side is the currently working AGM
and on the left side is the BSRS. All the instruments share the same light
source with the BSRT. 
\begin{figure}[!tbh]
    \centering
    \includegraphics[width=\columnwidth]{installation.jpg}
    \caption{installation picture}
    \label{fig:installation}
\end{figure}
The beam input is demonstrated by the yellow line. It shows beam arriving to
the MCP-PMT. On top of the reflective foil directing light towards the MCP-PMT
is a LED simulating the light source. This helps verifying the system
operation. The AGM amplifier is placed on top of the optical front-end to be
reachable in case of intervention. For reducing the light intensity, there are
pneumatically controlled optical filters in front of the PMT.

\section{Control Interface}
The gateware outputs data from four different block rams. The first stores raw
data from the ADC, that is used correctly set the initial delay of the
instrument's timing. It stores 1.5 turns and therefore it is published with
\SI{3.3}{Hz}. The next two memories store the actual snapshot data and averaged
data, but with finer granularity. They store 10 samples per \SI{10}{ms} window,
which gives the sample rate of \SI{100}{Hz}. The last memory is just bare
averaged data taken every \SI{10}{ms} defining the publishing rate. The
fetching handles the FESA server, which obtains interrupt every time the data
are ready. 
\begin{figure}[!tbh]
    \centering
    \includegraphics[width=\columnwidth]{control_interface.jpg}
    \caption{installation picture}
    \label{fig:control_interface}
\end{figure}
To simplify the setup tunining a new interface had to be created. It runs as
client of FESA server and displays the content of all the block rams. The main
control panel is depicted in Fig. \ref{fig:control_interface}. It displays
three plots, raw data from the ADC (top plot), processed averaged data (down
plot on the left) and and evolution of the data in time (down plot on the
right). The menu on the left controls the amplifier and bottom menu controls
the acquisition delays and optical filters.

\section{First acquired plots}
First measurements were taken in November 2023 on the ion beam. The
optical system splits the SL in ratio 1:8, so the new system gets
significantly less light then current one. In other word, we expect
lower value deviation for system measuring light with higher intensity.  
\begin{figure}[!htb]
    \begin{center}
        \scalebox{0.54}{\input{plots/ions.pgf}}
        \caption{Comparison of AGMs}
        \label{fig:comparison_chart}
    \end{center}
\end{figure}
The Fig. \ref{fig:comparison_chart} shows first acquisition of the new system
in comparison with the current one. The red plot is measured by the current
system and the blue chart is the new acquisition. In a range from 47th - 63th
bin, the gate is already close. There is just white noise, but the red plot
slowly decays and it continues below the zero line. This artefact is caused by
analog integrators and it is non-linear. The new system removes this effect.
The value deviation of the new acqusition from the current system represents
the yellow plot.   The relative value deviation of the new system from the
current one is depicted with the yellow plot. In this measurement the maximum
deviation is 4\% in the area of the decay of the red plot. The the deviation of
2\% in the steep parts can be caused by the sample phase alignment of the two
systems.

\section{Summary}
The AGM has been implemented on the VFC-HD carrier board simplifying
maintenance. The implementation is based on the direct digital conversion,
which eliminates artefacts in the output plot from the current system using
IBMS cards embedding analog integrators. It also solves the calibration hassle. 
The comparison of the two systems is summarized in the following table Tab.
\ref{tab:system_comparison}.
\input{tables/system_comparison.tex} 
The system is installed in the point 4 next to the current AGM collecting data
to be further analyzed. For the complet reliability test and proper function
verification, it needs to be collected more data and they have to be compared
with the current system. The systems shows so far, that with less light, it can
have still similar results to the current AGM.

\input{bibliography}
